Word Sense Disambiguation\\ \small{Leveraging the NLTK Toolkit and External Tools}

### 2011-04-14 paper/paper.tex

\documentclass{article}
\usepackage{url}
\usepackage{longtable}
\title{Word Sense Disambiguation\\
\small{Leveraging the NLTK Toolkit and External Tools}}
\author{
Alec Story
Ansu Abraham
Craig Frey\\
Dustin Tiedemann
Michael Zhu
Thomas Levine
Whitney Foster\\
}

@@

We  chose to implement our system in Python, and chose to use the machine
learning algorithms in the NLTK toolkit which is also Python-based.

We wrote our main control program, classify.py, so that it can choose either
[we're just doing nb and decision trees now right?] classifier and any
combination of  features. On top of this we wrote bash scripts that automated
looping through every combination of classifiers and features, along with
porting the results directly into the scorer.  Classify.py includes a help
switch -h, that explains the switches that control the different classifiers
and features along with any additional parameters that a certain feature might
support (for example, collocation has a window size parameter which defaults to
0) Performance was excellent for most of the features, allowing us to typically
run all of the words in the test set through every combination a classifier and
feature set in under an hour. (The number of runs can get quite high when you
factor in several parameters per feature, not just on/off)

@@

\section{Looking through Different Corpora}

In order to get more data to use with the bootstrapping algorithm, we attempted
to search through the Brown corpus and return all of the words that were being
disambiguated in the Senseval 3 task.  To do this we needed to find all of the
instances of the words that we needed to disambiguate and return a context
(list of words) and the position of the word in that list.  We also made sure
that we the context that we returned both started at the beginning of the
sentence and ended at the end of a sentence and had a minimum of 25 words
returned on either side of the word.  This data would then be converted into
the form of the data that we received from the Senseval document and ran
through the bootstrapping algorithm to have more data to train with.  However
when we actually tested this feature alone we found it to be very
uninformative.  First, there were very few instances of the words that we were
looking for (ex: activate only appeared 3 times), which made searching through
the entire corpus to find these words more time consuming than it was helpful.
Therefore, we decided not to use it for the actual test because it was not
informative enough by itself.

@@

%stepwise graph 1
%classifiert, collocation, cooccurrence  
%Interpret this graph

@@

%stepwise graph 2

@@

%stepwise graph 33

\subsection{Characterization of the top feature combinations}

%Let's get a comparison between our best performer(s) and the baseline

>>>>

\documentclass{article}
\usepackage{url}
\usepackage{longtable}
\usepackage{graphicx}
\title{Word Sense Disambiguation\\
\small{Leveraging the NLTK Toolkit and External Tools}}
\author{
Alec Story
Ansu Abraham
Craig Frey\\
Dustin Tiedemann
Michael Zhu
Thomas Levine
Whitney Foster\\
}

@@

We  chose to implement our system in Python, and chose to use the machine
learning algorithms in NLTK which is also Python-based.  We also used NLTK for its part of speech tagger, dependecy parser, and to import the Senseval data.  We coded the rest of the assignment on our own.

We wrote our main control program, classify.py, so that it can choose either
Naive Bayes or Decision Trees classifier and any 
combination of  features. On top of this we wrote bash scripts that automated
looping through every combination of classifiers and features, along with
porting the results directly into the scorer.  Classify.py includes a help
switch -h, that explains the switches that control the different classifiers
and features along with any additional parameters that a certain feature might
support (for example, collocation has a window size parameter which defaults to
0) Performance was excellent for most of the features, allowing us to typically
run all of the words in the test set through every combination a classifier and
feature set in under an hour. (The number of runs can get quite high when you
factor in several parameters per feature, not just on/off)

@@

\subsection{Looking through Different Corpora}

In order to get more data to use with the bootstrapping algorithm, we attempted
to search through the Brown corpus and return all of the words that were being
disambiguated in the Senseval 3 task.  To do this we needed to find all of the
instances of the words that we needed to disambiguate and return a context
(list of words) and the position of the word in that list.  We also made sure
that the context that we returned both started at the beginning of the
sentence and ended at the end of a sentence and had a minimum of 25 words
returned on either side of the word.  This data would then be converted into
the form of the data that we received from the Senseval document and ran
through the bootstrapping algorithm to have more data to train with.  However
when we actually tested this feature alone we found it to be very
uninformative.  First, there were very few instances of the words that we were
looking for (ex: activate only appeared 3 times), which made searching through
the entire corpus to find these words more time consuming than it was helpful.
Therefore, we decided not to use it for the actual test because it was not
informative enough by itself.

@@

\includegraphics[width=\textwidth]{pg_0001}
%stepwise graph 1
%classifiert, collocation, cooccurrence  
%Interpret this graph

@@

\includegraphics[width=\textwidth]{pg_0002}

@@

\includegraphics[width=\textwidth]{pg_0003}

\includegraphics[width=\textwidth]{pg_0004}
